Choosing an appropriate scalar transformation to normalise wpd

### 2021-02-04 scalar_transformation.Rmd
added transformation script

 - How the distribution of the raw wpd looks across nfacets and nx?

 Both shape and scale of the distribution changes for different nx and nfacet categories.

```{r read}
N01 <- read_rds("simulations/raw/null_design_quantrans/data-agg/all_data_wpd_N01.rds")
```

@@

 - Plot the values of wpd against nx*nfacet to see the rough relationship

```{r}
N01 %>% 
  ggplot(aes(x=as.factor(nx*nfacet), y = value)) +
  geom_boxplot() + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + xlab("nx*nfacet") + ylab("raw wpd")

```
 - Plot the values of wpd against nx*nfacet to see if the same relationship holds for different nx and nfacet relationships

Looks like for all the variations, there is a quadratic relationship between wpd values and nx*nfacet

```{r}
every_nth = function(n) {
  return(function(x) {x[c(TRUE, rep(FALSE, n - 1))]})
}
```

```{r}
N01 %>% 
  mutate(comp_lev = if_else(nx<nfacet, "nx<nfacet",
                            if_else(nx>nfacet,
                               "nx>nfacet", "nx=nfacet"))) %>% 
  ggplot(aes(x=as.factor(nx*nfacet), y = value)) +
  geom_boxplot() + facet_wrap(~comp_lev) + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + xlab("nx*nfacet") + ylab("raw wpd") + scale_x_discrete(breaks = every_nth(n = 5))
```

 - Attempt to linearize it with sqrt, sqrt(sqrt) and log (sqrt(sqrt))

 The transformation log (sqrt(sqrt())) on nx*nfacet finally makes it approximately linear.

```{r}
N01 %>% 
  mutate(comp_lev = if_else(nx<nfacet, "nx<nfacet",
                            if_else(nx>nfacet,
                                    "nx>nfacet", "nx=nfacet"))) %>% 
  ggplot(aes(x=sqrt(nx*nfacet), y = value)) +
  geom_point() + facet_wrap(~comp_lev) + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + ylab("raw wpd")

N01 %>% 
  mutate(comp_lev = if_else(nx<nfacet, "nx<nfacet",
                            if_else(nx>nfacet,
                                    "nx>nfacet", "nx=nfacet"))) %>% 
  ggplot(aes(x=sqrt(sqrt(nx*nfacet)), y = value)) +
  geom_point() + facet_wrap(~comp_lev) + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + ylab("raw wpd")

N01 %>% 
  mutate(comp_lev = if_else(nx<nfacet, "nx<nfacet",
                            if_else(nx>nfacet,
                                    "nx>nfacet", "nx=nfacet"))) %>% 
  ggplot(aes(x=log(sqrt(sqrt(nx*nfacet))), y = value)) +
  geom_point() + facet_wrap(~comp_lev) + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + ylab("raw wpd")

@@

 - Distribution after linearizing it

The distribution of $wpd_{norm} = log(sqrt(sqrt(nx*nfacet)))/wpd$ is plotted. The shape and spread look similar but location is shifting to the right.

(If we define $wpd_{norm}$ as the inverse of it, the values become too small and the distribution too skewed. Hence the inverse of it is considered.)

@@

```{r}
 N01 %>% 
  ggplot(aes(x = forecast::BoxCox(log(sqrt(sqrt(nx*nfacet)))/value,
             lambda = "auto"))) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("wpd norm")

>>>>

 - How the distribution of the raw wpd (without any transformation for normalization) looks across nfacets and nx? Both shape and scale of the distribution changes for different nx and nfacet categories.

```{r read}
N01 <- read_csv("simulations/raw/null_design_quantrans/data-agg/all_data_wpd_N01.csv")
```

@@

 - Plot the values of raw wpd against nx*nfacet to see the rough relationship

```{r}
N01 %>% 
  ggplot(aes(x=nx*nfacet, y = value)) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue") + xlab("nx*nfacet") + ylab("raw wpd")

@@

**Current attempt**

- Fit a log-linear relationship of median(values) to log(nx*nfacet) as discussed today. The fit is pretty good.

```{r, echo = TRUE}
N01_median <- N01 %>% 
  group_by(nx*nfacet) %>% 
  summarise(actual = median(value))

fit_lm2 <- lm(actual ~ poly(log(`nx * nfacet`) ,1, raw=TRUE), data = N01_median)
summary(fit_lm2)
```

- Make it a horizontal line so that the values are not affected by 
nx\*nfacet by plotting (value - intercept)/log(nx\*nfacet) against log(nx*nfacet). The line actually becomes horizontal after this transformation.

```{r}
intercept <- fit_lm2$coefficients[1]
slope <- fit_lm2$coefficients[2]

N01 %>% 
  ggplot(aes(x=log(nx*nfacet), y = (value - intercept)/log(nx*nfacet))) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue")

```
- See distribution of the transformed variable across nx and nfacets. For higher values of nx and nfacet, the distribution is similar, but not for lower values.

```{r}
N01 %>% 
  ggplot(aes(x = (value - intercept)/(log(nx*nfacet)))) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("wpd") +
  scale_x_continuous(breaks = scales::breaks_extended(3))

**Earlier attempt**
 - The distribution of $wpd_{norm} = log(sqrt(sqrt(nx*nfacet)))/wpd$ is plotted since the plot of log(sqrt(sqrt(nx * nfacet))) against wpd looked linear. The shape and spread look similar but location is shifting to the right. (If we define $wpd_{norm}$ as the inverse of it, the values become too small and the distribution too skewed. Hence the inverse of it is considered.)

@@

```{r}

N01 %>% 
  ggplot(aes(x=log(nx*nfacet),
             y = log(sqrt(nx*nfacet))/value)) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue")

 N01 %>% 
  ggplot(aes(x = log(sqrt(nx*nfacet))/value)) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("wpd norm")
```

```{r}
fit2 <- lm(log(sqrt(nx*nfacet))/value ~ log(nx*nfacet), N01)
intercept <- fit2$coefficients[1]
slope <- fit2$coefficients[2]

N01 %>% 
ggplot(aes(x = 
             (log(sqrt(nx*nfacet)) - intercept)/log(nx*nfacet))) +
geom_density(fill = "blue") +
facet_grid(nx~nfacet,
labeller = "label_both") +
xlab("wpd norm")
